Update and evaluation of the ozone dry deposition in the Oslo~CTM3 v1.0

### 2019-07-10 DryDep.tex
Finished final revision of text. ToDo - diff and new abstract pdf.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%      
\section{Evaluation}
\label{sec:eval}
In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the mOSaic scheme. We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, ozone concentrations in the lowermost model level $[\chem{O_3}](p_0)$, and tropospheric burden $\sum_\text{trop}\chem{O_3}$. We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}), the MACC-reanalysis (Section~\ref{subsec:macc}), and observations (Section~\ref{subsec:obs}). The Oslo~CTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4). Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}, while emissions from soil and wetlands are computed by MEGAN. Resultant $\mathrm{NO_x}$ emissions are up-scaled to match Global Emissions InitiAtive (GEIA) inventory. For oceanic emissions of \chem{CO}, we use predefined global fields from POET \citep{POET}. Emissions of \chem{CH_4} are taken from the EU project (EU GOCE 037048) Hydrogen, Methane and Nitrous oxide: Trend variability, budgets and interactions with the biosphere (HYMN) for the year 2003 and scaled to oceanic amounts of $\mathrm{CH_4}$ from NASA. In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sensitivity studies}
\label{subsec:sens}
Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every single difference in results to a respective change. Therefore, we conducted one reference simulation denoted as \emph{mOSaic} implementing all parameters as tabulated in \citet{ACP:Simpson2012} and in total seven sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}). A reference simulation featuring the Oslo~CTM3 Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
All model experiments discussed in the following are summarized in Table~\ref{tab:simsum}. An \emph{x} therein denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.
For all model integrations, the meteorological reference year is 2005. This choice affects the direct comparison with data and studies that either show results based on decadal averages or differing years.

@@

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $R^\chem{O_3}$ & Emissions\\
    & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & & (year)\\
    \middlehline
    Wesely\_type      &  \multicolumn{3}{c}{n/a}   & x & n/a& \citet{AE:Wesely1989, JGR:Hough1991}  & 2005 \\
    mOSaic            &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offLight  &   x & 1 & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offPhen   &   1 & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_SWVL1     &   x & x & VL1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_ice       &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
    mOSaic\_desert    &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{AE:Gusten1995}$^*$  & 2005 \\
    mOSaic\_emis2014  &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2014 \\
    mOSaic\_hough     &   x & x & x                &   &  x & \citet{AE:Wesely1989, JGR:Hough1991}$^\star$  & 2005 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.7} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig04}
  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average ozone burden in the lowermost model level; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition. The ozone burden increases globally except for some regions covered by tropical forest. Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to this divergence especially in desert regions is the difference between the prescribed surface resistances $R^\chem{O_3}$ in the Wesely scheme in comparison to those used in \emph{mOSaic}. We cover this in more detail in the following section.

@@

Dry deposition velocities are directly available only for the new model version. For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(z_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$ 
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
  \label{eq:retro_vdd}
\end{equation}
Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$. In case of \emph{mOSaic}, Eq.~(\ref{eq:retro_vdd}) yields similar values for $v^\chem{O_3}_\text{DD}$ as directly available from model output.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\mathcal{O}(5\,\unit{ppb})$). We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies focusing on the stomatal conductance, there is only a low absolute variance. Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly. Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}) amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$). Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{mOSaic\_ice} displays a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} ($\mathcal{O}(20\,\unit{ppb})$) but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Reducing $R^\chem{O_3}_\text{desert}$ by $60\,\unit{\%}$ (\emph{mOSaic\_desert}) a reduction in the order of $1\,\unit{ppb}$ is found mainly limited to the NH. The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.7). The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{mOSaic} compared to \emph{mOSaic\_emis2014}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, where an increase in ozone precursors is seen in CEDS.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b. The dry deposition velocities in the mOSaic scheme are well below the Wesely scheme and in remarkable agreement with the results shown by \citet{ACP:Hardacre2015}. In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean. This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} dry deposition velocity above ice and snow should be considered as new standard for the Oslo~CTM3. This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation. The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$. The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.6: Fig.~S8). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c. In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies. In the most extreme case (NH subtropics and mid latitudes), the total ozone dry deposition drops to one-half of the amount given by \emph{Wesely\_type}. The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a substantial difference in $v_\text{desert}^\chem{O_3}$. Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the mOSaic scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameter set \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}. These differences in surface dry deposition velocities over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}). We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).

There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme match better. This could be a sign of differences in photo-chemistry and transport (e.g. convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further. Another reason might be found within the (chemistry) routines in the Oslo~CTM3 that actually compute the removal of ozone from the atmosphere which might have been adjusted to the less physical dry deposition velocities in the past but this is subject to further investigations.

@@

As suggested by \citet{ACP:Hardacre2015}, we also look at ozone dry deposition velocities with respect to surface types separately. Since dry deposition velocities are not directly available for \emph{Wesely\_type}, we use Eq.~(\ref{eq:retro_vdd}) to calculate these. Based on a CLM~2 average dynamic land surface map, we generate masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and use these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest. Thus, it is not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type. Since we have not performed a full unfolding on the data, the results should be treated with slight caution (e.g. over cropland). In case of the mOSaic scheme, we have pre-selected the dry deposition velocities in accordance to the land surface type.
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{fig06}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}
\end{figure*}
%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%      
\section{Evaluation}
\label{sec:eval}
In this section, we present results from a manifold of Oslo~CTM3 model integrations testing different parameters of the mOSaic scheme. We focus on changes in ozone total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, concentrations in the lowermost model level $[\chem{O_3}](p_0)$, and tropospheric burden $\sum_\text{trop}\chem{O_3}$. We evaluate our results with respect to the multi-model comparison of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}), the MACC-reanalysis (Section~\ref{subsec:macc}), and observations (Section~\ref{subsec:obs}). The Oslo~CTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4). Biogenic emissions are taken from MEGAN-MACC output \citep{ACP:Sindelarova2014}, while emissions from soil and wetlands are computed by MEGAN. Resultant $\mathrm{NO_x}$ emissions are up-scaled to match Global Emissions InitiAtive (GEIA) inventory. For oceanic emissions of \chem{CO}, we use predefined global fields from POET \citep{POET}. Emissions of \chem{CH_4} are taken from the EU project (EU GOCE 037048) Hydrogen, Methane and Nitrous oxide: Trend variability, budgets and interactions with the biosphere (HYMN) for the year 2003 and scaled to oceanic amounts of $\mathrm{CH_4}$ from NASA. In the following (Section~\ref{subsec:sens}), we will present the various model sensitivity studies.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sensitivity studies}
\label{subsec:sens}
Due to significant differences between the mOSaic scheme and the previous Wesely scheme with respect to implementation, it is not possible to fully disentangle and trace back every single difference in results to a respective change. Therefore, we conducted one reference simulation denoted as \emph{mOSaic} and in total seven sensitivity studies to probe the parameter space for stomatal conductance (\emph{mOSaic\_offLight}, \emph{mOSaic\_offPhen}, and \emph{mOSaic\_SWVL1}), ozone surface resistance $R^\chem{O_3}$ (\emph{mOSaic\_ice}, \emph{mOSaic\_desert}, and \emph{mOSaic\_hough}), and emissions (\emph{mOSaic\_emis2014}). A reference simulation featuring the Oslo~CTM3 Wesely scheme has been conducted and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models.
All model experiments discussed in the following are summarized in Table~\ref{tab:simsum}. An \emph{x} therein denotes that the model was run exactly in the configuration and with parameters as has been described in Section~\ref{sec:model_des}.
For all model integrations, the meteorological reference year is 2005. This choice affects the direct comparison with data and studies that either show results based on decadal averages or differing years, because non-linearities in ozone formation and destruction make ozone concentrations sensitive to both, differences in local concentration of precursors and meteorological conditions \citep{JGR:Jin2013}.

@@

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $R^\chem{O_3}$ & Emissions\\
    & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & & (year)\\
    \middlehline
    Wesely\_type      &  \multicolumn{3}{c}{n/a}   & x & n/a& \citet{AE:Wesely1989, JGR:Hough1991}  & 2005 \\
    mOSaic            &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offLight  &   x & 1 & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offPhen   &   1 & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_SWVL1     &   x & x & VL1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_ice       &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
    mOSaic\_desert    &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{AE:Gusten1995}$^*$  & 2005 \\
    mOSaic\_emis2014  &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2014 \\
    mOSaic\_hough     &   x & x & x                &   &  x & \citet{AE:Wesely1989, JGR:Hough1991}$^\star$  & 2005 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.7} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig04}
  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average ozone burden in the lowermost model level; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition. The ozone burden increases globally except for some regions covered by tropical forest. Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to this divergence especially in desert regions is the difference between the prescribed surface resistances $R^\chem{O_3}$ in the Wesely scheme in comparison to those used in \emph{mOSaic}. We come back to this in the following sections.

@@

Dry deposition velocities are directly available only for the new model version. For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(p_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$ 
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(p_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
  \label{eq:retro_vdd}
\end{equation}
Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$. In case of \emph{mOSaic}, resulting values for $v^\chem{O_3}_\text{DD}$ from Eq.~(\ref{eq:retro_vdd}) are compatible with the values which are directly ailable from model output.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\mathcal{O}(5\,\unit{ppb})$). We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies focusing on the stomatal conductance, there is only a low absolute variance. Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly. Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}) amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$). Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{mOSaic\_ice} displays a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} ($\mathcal{O}(20\,\unit{ppb})$) but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Reducing $R^\chem{O_3}_\text{desert}$ by $60\,\unit{\%}$ (\emph{mOSaic\_desert}) a reduction in the order of $1\,\unit{ppb}$ is found mainly limited to the NH. The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.7). The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{mOSaic} compared to \emph{mOSaic\_emis2014}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, where an increase in ozone precursors is seen in CEDS.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b. The dry deposition velocities in the mOSaic scheme are well below the Wesely scheme and in remarkable agreement with the results shown by \citet{ACP:Hardacre2015}. In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean. This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} surface resistance above ice and snow should be considered as new standard for the Oslo~CTM3. This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation. The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$. The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.6: Fig.~S8). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$ (almost entirely covered by ocean), $(12-30)\,\unit{^\circ S}$ (coinciding with the location of Australia and its desert regions), as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c. In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies. In the most extreme case (NH subtropics and mid latitudes), the total ozone dry deposition drops to one-half of the amount given by \emph{Wesely\_type}. The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a substantial difference in $v_\text{desert}^\chem{O_3}$. Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the mOSaic scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original parameter set \citep[$v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$,][]{ACP:Simpson2012}. These differences in surface resistances over huge parts of the unvegetated surface of the Earth accounts for most of the qualitative difference between the Wesely and the mOSaic scheme, but does not explain the quantitative difference (compare \emph{mOSaic\_hough}). We further elaborate on this in the following (Section~\ref{subsubsec:seasons}).

There seems to be a discrepancy between the Oslo~CTM3 response and the multi-model-mean, since the Wesely scheme is similar to the multi-model-mean with respect to total annual ozone dry deposition, while the $v_\text{DD}^\chem{O_3}$ of the mOSaic scheme match better. This could be a sign of differences in photo-chemistry and transport (e.g. convective, advective, STE) between the Oslo~CTM3 and the average TF~HTAP model, but without comparing to the actual $[\chem{O_3}]$ of the TF~HTAP models that participated in the model intercomparison, we cannot elaborate on this any further. This may also hint to issues in the Oslo~CTM3 photo-chemistry, which may have a too high ozone production, or the actual removal of ozone from the atmosphere, which might have been adjusted to the less physical dry deposition velocities in the past, but this is subject to further investigations.

@@

As suggested by \citet{ACP:Hardacre2015}, we also look at ozone dry deposition velocities with respect to surface types separately. Since dry deposition velocities are not directly available for \emph{Wesely\_type}, we use Eq.~(\ref{eq:retro_vdd}) to estimate these. Based on a CLM~2 average dynamic land surface map, we generate masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and use these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest. Thus, it is not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type. Since we have not performed a full unfolding on the data, the results should be treated with slight caution (e.g. over cropland). In case of the mOSaic scheme, we have pre-selected the dry deposition velocities in accordance to the land surface type.
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{fig06}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}
\end{figure*}
%

In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the mOSaic scheme and the multi-model-mean. The divergence of the average dry deposition velocities between \emph{Wesely\_type} and \emph{mOSaic} in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.20\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.65\,\unit{cm\,s^{-1}}$) is quite remarkable. The difference of \emph{mOSaic} and \emph{Wesely\_type} to the multi-model-mean in tropical forest regions is $\Delta_\text{mOSaic} \overline{v}_\text{tropical forest}^\chem{O_3} = 0.61\,\unit{cm\,s^{-1}}$ and $\Delta_\text{Wesely\_type} \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$, respectively. The multi-model-mean displays a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$), which cannot be reproduced with the mOSaic scheme. The dry deposition velocities over desert regions are consistent with the average values from the prescribed ozone surface resistances, which means that in the mOSaic scheme they are $1$ order of magnitude lower than in the Wesely scheme. In the mOSaic scheme, dry deposition to deserts is dominated by contribution from $R_b$. From a limited number of ozone flux measurements in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the mOSaic scheme. Similarly, the dry deposition velocities over water differ. From measurements during ship-campaigns a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ over the ocean has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the mOSaic scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean, but is in line with the multi-model-mean. This implies that ozone concentrations might even become larger and dry deposition even lower in the model if a more advanced dry deposition scheme to the ocean would be implemented. With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs and phenologies, especially in regions covered by tropical forest \citep{GCB:Anav2017} or in boreal regions. 
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{fig07}
  \caption{Average seasonal cycles of ozone dry deposition velocities separated by land use type. Results from \citep{ACP:Hardacre2015} are shown as a reference. We refrain from showing the full extent of \emph{EMEP\_offLight} here, since it is an extreme scenario and has been discussed already.}
  \label{fig:mmm_drydep_season_pft}
\end{figure*}
%
